---
title: "Top Industries in SLO County Based on Location Quotients"
format: pdf
editor: visual
---

```{r, echo=FALSE, eval=FALSE}
install.packages("readr")
install.packages("readxl")
install.packages("dplyr")
install.packages("stringr")
install.packages("tidyr")
```

```{r, echo=FALSE, eval=FALSE}
library(readr)
library(readxl)
library(dplyr)
library(stringr)
library(tidyr)
```

# Using Location Quotients (LQs) to identify important industries within San Luis Obispo (SLO) County.

This project explore the industrial composition of San Luis Obispo (SLO) County by using **Location Quotients (LQs)** to identify industries that are more concentrated locally compared to the national economy. The purpose is to determine which sub-sectors and industries play a disproportionately significant role in SLO County's economy.

## 1. Loading and Preparing the Data

To begin this process, we loaded three data sets:

-   `naics_codes.xlsx`: provides NAICS industry codes and their titles

    -   This `.xlsx` file is a bit tricky because of the fact there are unnecessary details and format quirks so I had to do a bit of cleaning in order to get codes and titles.

-   `slo_employment.csv`: contains monthly employment levels by industry in SLO County

-   `us_employment.csv`: contains comparable national data

```{r}
slo <- read_csv("slo_employment.csv")
us <- read_csv("us_employment.csv")
naics <- read_excel("naics_codes.xlsx", skip = 2) %>%
  select(2:3) %>%
  rename(codes = 1, title = 2) %>%
  mutate(title = str_remove(title, "T$")) %>%
  drop_na()
```

## 2. Summarizing Employment by Industry

Next, we aggregated employment data fro each industry at the annual level by summing the employment figures for each `industry_code.`

```{r}
slo_sum <- slo %>%
  group_by(industry_code) %>%
  summarise(emp_slo = sum(month1_emplvl, na.rm = TRUE))

us_sum <- us %>%
  group_by(industry_code) %>%
  summarise(emp_us = sum(month1_emplvl, na.rm = TRUE))
```

## 3. Calculating Location Quotients (LQs)

We merged the SLO and US employment summaries by industry code and then computed the location quotient:

$$
LQ_i = \left( \frac{e^{\text{SLO}}_i}{e^{\text{SLO}}} \right) \Big/ \left( \frac{e^{\text{US}}_i}{e^{\text{US}}} \right)
$$

Where:\
- ( e\^{\text{SLO}}\_i ) = Employment in industry *i* in **SLO**\
- ( e\^{\text{SLO}} ) = Total employment in **SLO**\
- ( e\^{\text{US}}\_i ) = Employment in industry *i* in **US**\
- ( e\^{\text{US}} ) = Total employment in **US**

```{r}
lq_data <- slo_sum %>%
  inner_join(us_sum, by = "industry_code") %>%
  mutate(share_slo = emp_slo / sum(emp_slo),
         share_us = emp_us / sum(emp_us),
         LQ = share_slo / share_us)
```

The LQ tells us whether a given industry is **overrepresented (LQ \> 1)** or **underrepresented (LQ \< 1)** in SLO compared to the national average.

## 4. Identifying Top 5 NAICS Sub-Sectors

We focused on **3-digit NAICS codes**, which represent broad sub-sectors within the economy (e.g. Food Services, Construction). Filtering and sorting by highest LQs helped us identify the **top 10 most concentrated sub-sectors** in SLO County.

```{r}
mod_1_lq_data <- lq_data %>%
  mutate(code_str = as.character(industry_code)) %>%
  filter(nchar(code_str) == 3) %>%
  select(industry_code, LQ, code_str) %>%
  arrange(desc(LQ)) %>%
  slice_max(LQ, n = 5) %>%
  print()
```

## 5. Drilling Down: Top Industries Within Each Sub-Sector

For each of the top 5 sub-sectors, we identified the **top 5 detailed industries** (based on full-length NAICS codes) with the highest LQs. This deeper look gives us a finer-grained understanding of what specific industries drive the sub-sector's prominence.

```{r}
mod_2_lq_data <- lq_data %>%
  mutate(code_str = as.character(industry_code)) %>%
  mutate(naics_3digit = substr(industry_code, 1, 3)) %>%
  group_by(naics_3digit) %>%
  filter(naics_3digit %in% mod_1_lq_data$code_str) %>%
  select(industry_code, LQ, naics_3digit) %>%
  arrange(desc(LQ)) %>%
  slice_max(LQ, n = 5) %>%
  print()
```
